---
title: "In Class Notes"
author: "Cameron Smith"
date: 'Last updated: `r format(Sys.time(), "%b %d, %Y - %I:%M %p")`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#10/9/2018

## Heads or Tails

Let x = number of heads in n = 10 flips of a fair coin.

x could be 0, 1, 2, 3, 4, 5, 6, 7, 8, 9, 10.

P(x = 0) = $(1/2)^10$ = T, T, T, T, T, T, T, T, T, T

P(x = 1) = $(1/2)^9(1/2)10$ = H, T, T, T, T, T, T, T, T, T

P(x = 2) = $(1/2)^8(1/2)^245$ = H, H, T, T, T, T, T, T, T, T

###Use `choose()`.

```{r}
choose(10,0:10)
```

```{r}
plot(x = 0:10, y = dbinom(0:10, 10, 1/2), type = "h", xlab = "x", ylab = expression(P(X == x)))

x <- 0:10
px <- dbinom(x, 10, 1/2)
EX <- sum(x*px)
EX

VX <- sum((x - EX)^2 * px)
VX

```

#10/10/2018

$Z = \frac{stat - M_{stat}}{\theta_{stat}}$

$t = \frac{stat - M_{stat}}{\widehat\theta_{stat}}$

If $X\sim(M,\theta)$ provided n is sufficently large

$\bar{X} \sim N(\mu_{\bar{x}}=\mu, \theta_{\bar{x}} = \frac{\theta}{\sqrt{n}})$

```{r}
sims <- 10000
xbar <- numeric(sims)
for (i in 1:sims) {
  x <- runif(10, 0, 10)
  xbar[i] <- mean(x)
}

plot(xbar)
mean(xbar)
var(xbar)
sqrt(100/12)/4
```

$\mathbb{P}(Z_{\frac{\sigma}{Z}}\le Z \le Z_{1-\frac{\sigma}{Z}}) = 1 - \sigma$

$\mathbb{P}(Z_{\frac{\sigma}{Z}}\le \frac{\bar{x} - \mu}{\theta/\sqrt{n} } \le Z_{1-\frac{\sigma}{Z}}) = 1 - \sigma$