---
title: "RMarkdown Examples"
author: "Cameron Smith"
date: "11/13/2018"
output: bookdown::html_document2
---

```{r setup, include=FALSE, echo = TRUE, fig.align='center', warning=FALSE, message=FALSE, comment= NA}
knitr::opts_chunk$set(echo = TRUE)
```
# R Markdown

When you click the **Knit** button a dosument will be generated that includes both content as well as the output of any embedded R code chunks within the document.  Consider Table \@ref(tab:cars).

```{r cars, echo=FALSE, fig.cap="First 6 rows of mtcars"}
knitr::kable(head(mtcars), booktabs = TRUE, caption = "First 6 rows of `mtcars`")
```

# Including Plots
You can also embed plots such as Figure \@ref(fig:pressure)

```{r pressure, echo = FALSE, fig.cap="Scatterplot of `pressure` versus `temperature`"}
plot(pressure)
```

#3

# Chi-Square Test?

Two different professors teach and introductory statistics course. Table \@ref(tab:TP) shows the distribution of final grades they reported.  We wonder whether one of these professors is an "easier" grader.


```{r TP, echo = FALSE}
mat <- matrix(c(3, 10, 11, 11, 14, 8, 10, 1, 4, 0), nrow = 5, byrow=TRUE)
dimnames(mat) <- list(Grade = c(LETTERS[1:4], LETTERS[6]), Professor = c("Alpha", "Beta"))
tt <- as.table(mat)
knitr::kable(tt, caption = "Grades for Alpha and Beta")
```

If you reject the null hypothesis, the only error you can make is a Type I error.
If you fail to reject the null hypothesis you can only make a Type II error.
